Parity games in pushdown systems with pebbles

### 2023-06-11 reachability/article.tex
up

\renewcommand{\A}{\mathcal{A}}
\newcommand{\B}{\mathcal{B}}
\renewcommand{\C}{\mathcal{C}}
\newcommand{\T}{\mathcal{T}}
\renewcommand{\G}{\mathcal{G}}
\renewcommand{\O}{\mathcal{O}}
\renewcommand{\R}{\mathbb{R}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\Time}{\mathbb{T}}
\newcommand{\Const}{\mathsf{Consts}}
\newcommand{\PTA}{\textsc{PTA}}
\newcommand{\MSO}{\textsc{MSO}}

@@

@@

>>>>

\renewcommand{\A}{\mathcal{A}}
\newcommand{\B}{\mathcal{B}}
\renewcommand{\C}{\mathcal{C}}
\newcommand{\T}{\mathcal{T}}
\renewcommand{\G}{\mathcal{G}}
\renewcommand{\O}{\mathcal{O}}
\renewcommand{\R}{\mathbb{R}}
\renewcommand{\H}{\mathcal{H}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\Time}{\mathbb{T}}
\newcommand{\Const}{\mathsf{Consts}}
\newcommand{\Conf}{\mathsf{Conf}}
\newcommand{\PTA}{\textsc{PTA}}
\newcommand{\MSO}{\textsc{MSO}}

@@

\newcommand{\Op}{\mathsf{Op}}
\newcommand{\win}{\textsc{Win}}

@@

%\subsection{A decidability result from reduction to parity games on automata with pebbles}\label{pebbles}
\section{%A decidability result via r
Reduction to pebble pushdown automata parity games
}\label{pebbles}

% \mh{Motivates the use of pebbles. See pebbles on two-way automata, two-way tree automata, weighted automata, etc.}

% 21. Globerman, N., Harel, D.: Complexity results for two-way and multi-pebble automata and their logics. Theor. Comput. Sci. 169, 161–184 (1996)
%		same as [13] bellow
%
% Engelfriet, J., Hoogeboom, H.J.: Tree-walking pebble automata. In: Karhumäki, J., Maurer, H., Pa ̆un, G., Rozenberg, G. (Eds.) Jewels are Forever, Contributions to Theoretical Computer Science in Honor of Arto Salomaa, pp. 72–83. Springer, Berlin (1999) for tree-walking automata.
%
% @incollection{engelfriet1999tree,
%  title={Tree-walking pebble automata},
%  author={Engelfriet, Joost and Hoogeboom, Hendrik Jan},
%  booktitle={Jewels are forever},
%  pages={72--83},
%  year={1999},
%  publisher={Springer}
%}
% 
%
% Two-way pebble transducers for partial functions and their composition, Joost Engelfriet
%
% Tree-Walking Pebble automata, Joost Engelfriet and Hendrik Jan Hoogeboom 
A pebble automaton is a two-way finite state automaton that uses a fixed, finite number of
pebbles that it can drop on, and lift from 
words, using them as markers.
Pebble automata recognize regular languages only, provided the life times of the pebbles, i.e. the times between dropping a pebble and lifting it again, are properly nested
\cite{globerman1996complexity, engelfriet1999tree}.
Automata with nested pebbles were also introduced
for tree-walking automata. 
It is known that tree-walking automata do not recognize all regular tree languages \cite{bojanczyk2008tree}%, the main trouble being that tree-walking automata get lost rather easily
.
Using pebbles is a remedy against getting lost along a tree, but 
the unrestricted use of pebbles leads to a class of tree languages much larger than the
regular tree languages, in fact to all tree languages in $\NSPACE(log ~ n)$.
Thus, in  both pebble word automata and pebble tree-walking automata, the placement of the pebble follows a strict stack discipline. It is traditional hence to represent syntactically the dropping and lifting of pebbles by operations \textit{lift} and \textit{drop}; a \textit{drop} simply records the current position with a fresh pebble (such a pebble should be available) % and returns to the beginning of the tape, 
and a \textit{lift} pops the last dropped pebble if the current position corresponds to the one recorded by it.

In this section, we extend these ideas to pushdown automata. Instead of using pebbles as markings on their input, pebble pushdown automata have the ability to lift or drop pebbles on their universe of stack contents; a \textit{drop} simply records the current stack content with a fresh pebble (such a pebble should be available) % and returns to the beginning of the tape, 
and a \textit{lift} pops the last dropped pebble 
while requiring that
% assuming
 it was placed on the current node. 
% Thus hierarchicality on the registers works slightly differently from hierarchicality on the pebbles, hierarchicality on the registers being a syntactic property, whereas hierarchicality on the pebbles is a semantical one.
One can think of a pebble as a register that can store a stack content for later comparisons.

We first define more formally our pebble pushdown automaton framework, and then provide a reduction from the problem of solving parametric pushdown parity games to the problem of solving pebble pushdown parity games.

%\subsubsection{Pebble pushdown automata}
\subsection{Pebble pushdown automata}

\newcommand{\ppda}{\mathcal{I}}
% \Denarius
% \vernal
% \mathscr{Z}
% \mathcal{I}
% \pluto
% \capricornus
% \Bumpeq
% \Pfund

An {\em $n$-pebble pushdown automaton} 
is a tuple 
$\ppda = (Q, \Gamma,  R, q_{init}, \gamma_{init}, F)$
where
\begin{itemize}
\item $Q$ is a {\em finite set of  states},
\item $\Gamma$ is a {\em finite stack alphabet},
\item  $R  \subseteq  Q  \times \Gamma \times
		 \{ 0, 1, \ldots n\} \times \mathcal{P}(\{1,\ldots,n\})
		 \times Q  \times (\Op(\Gamma) \cup \{\text{\textit{drop}}, \text{\textit{lift}}\})$ is a {\em finite set of rules},
		 where the fourth element $S$ of a rule $r$ is a subset of 
		$\mathcal{P}(\{1,\ldots,i\})$ where $i$ is the third element of $r$,
		and if the last element is
		$\text{\textit{lift}}$ then we additionally require $i \in S$,
\item $q_{init} \in Q$ is an {\em initial  state}, 
\item $\gamma_{init} \in \Gamma$ is an {\em initial stack symbol}, and
\item $F \subseteq Q$ is a {\em set of final  states}.
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
% \mh{remind the reader what a partial function is}

\par\noindent\ignorespacesafterend 
Recall that for \iffalse a partial function $f$ from a set $S$ to a set $X$ 
%is a function defined on a subset $C$ of $S$ (possibly $S$ itself) with output values in $X$, and 
that is defined on a subset $C \subseteq S$
is denoted by
$ f : S \rightharpoonup X $.

For \fi a partial function $ f : S \rightharpoonup X $ (see page~\pageref{partial}), for notational purposes, we consider some element $\bot_X \not\in X$ and 
 associate $f$ with the function returning the 
bottom element $\bot_X$ when $f$ is undefined. Thus we write $(X \biguplus \{ \bot_X \})^S$ for the set of all partial functions from $S$ to $X$, which we here abbreviate as $(X \biguplus \{ \bot \})^S$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% An $i$-configuration of $\ppda$ is an element of
%$ Q\times W^* \times (W^*)^{\{1, \ldots i\}}$.
% By $\Conf_i(\ppda)$ we denote the set of $i$-configurations of $\ppda$.
By $\Conf(\ppda)=Q\times \Gamma^* \times (\Gamma^* \biguplus \{ \bot \})^{\{1, \ldots, n\}}$ we denote the set of
% By $\Conf(\ppda) = \bigcup_{0 \leq i \leq n} \Conf_i(\ppda)$ we denote the set of
{\em configurations} of $\ppda$. As expected, we rather write $q(w, \mu)$ instead of $(q, w, \mu)$.
An {\em$i$-configuration} for $i > 0$ of $\ppda$ is a configuration $q(z,\mu)$ where
$\text{Dom}(\mu) = \{1, \ldots, i\}$, while a $0$-configuration is a configuration 
 $q(z,\mu)$ where $\text{Dom}(\mu)=\emptyset$.

The idea of a transition $(q, a, i, S, q', m) \in R$
is that, if the automaton $\ppda$ is in state $q$ with pebbles $1,\ldots, i$ dropped \---- or without pebble dropped if $i = 0$ \---- with top stack symbol $a$, and stack content which corresponds to the stack contents of the pebbles from $S$ and only these pebbles, then
$\ppda$ goes to state 
$q'$ and makes modifications to the stack or the pebbles according to
$m$. 
% \mh{The set $S$ is used for testing the presence of certain pebbles.}
Note a pebble 
can be lifted only if the stack content which corresponds to the pebble
is the same as the current stack content. 
%$\Dzeta$ $\zeta$ $\Xi$
This is enforced by syntactically requiring
the last pebble dropped $i$ is in the set $S$ used for testing the presence of certain pebbles.

\iffalse
A {\em weak} $n$-pebble pushdown system is an $n$-pebble pushdown system $\ppda = (Q, \Gamma,  \Delta )$ in which $\text{\textit{lift}}$ operations also require the presence of the last pebble dropped, i.e.
for all $(q, a, i, S, q', \text{\textit{lift}}) \in \Delta$, we have $ i \in S $. 
%  is an additionnal restriction $f(i) = wa$ that needs to hold for the application of the $lift$ operation.
{\em Weak $n$-pebble pushdown games} are defined naturally as $n$-pebble pushdown games on 
weak $n$-pebble pushdown systems.

\fi

A {\em pebble set} of $\ppda$ is a set $U \subseteq \{ 1, \ldots, n\}$. For a stack alphabet $\Gamma$, a 
{\em $U \!$-pebble assignment} is a function which maps each $j \in U$ to a word in $ \Gamma^*$.
The $\emptyset$-pebble assignment is denoted by $\mu_{init} :  \{ 1, \ldots, n\} \rightharpoonup  \Gamma^* $ and is the 
%partial function from  $\{ 1, \ldots, n\}$ to $ \Gamma^* $ that
%does not map any $j \in \{ 1, \ldots, n \}$ to an output.
totally undefined function.

\iffalse
For $0 \leq i \leq n$, an {\em $i$-configuration} is a tuple $(q, w, \mu )$, where $w \in \Gamma^*$, $q$ is a state and
$\mu$ a $\{ 1, \ldots , i \}$-pebble assignment. 
We call $w$ the current stack, $q$ the current state and
$\mu$ the current pebble assignment. 
We also write $(q,w, w_{1} , \ldots , w_{i} )$ if 
$\mu(j) = w_j$, for each
$j \leq i$.

The set of configurations is the union of all $i$-configurations for $0 \leq i \leq n$. 
\fi

\begin{samepage}
%\begin{definition}
An $n$-pebble pushdown automaton $\ppda =  (Q, \Gamma,  R, q_{init}, {\gamma_{init}} , F)$ induces 
a transition system  $T_{\ppda} = (\Conf(\ppda), \rightarrow_{\ppda})$ 
 such that
for all $(q, a, i, S, q', m) \in R$, with $a \in \Gamma$,
 for all words $w \in  \Gamma^*$, and for all $\{ 1, \ldots , i \}$-pebble assignments $\mu$ such that $\mu(j) = wa$ 
 %(respectively $\mu(j)= {\gamma_{init}}  $) 
 for each
$j \in S$ 	and
 $\mu(j) \neq wa$ 
 %(resp. $\mu(j) \neq {\gamma_{init}} $) 
 for each $j \in \{ 1, \ldots , i \} \setminus S$,
the following holds
\begin{itemize}
\item if $m \in \Op(\Gamma)$, 
% $(q,wa,\mu) \rightarrow_{\ppda} (q',wm, \mu)$, 
either
\begin{itemize}
\item $ m = \text{\textit{push}}^\gamma$ and $(q,wa,\mu) \rightarrow_{\ppda} (q',wa\gamma, \mu)$,

\item $ m = \text{\textit{pop}}$ and $(q,wa,\mu) \rightarrow_{\ppda} (q',w, \mu)$, or

\item $ m = \text{\textit{skip}}$ and $(q,wa,\mu) \rightarrow_{\ppda} (q',wa, \mu)$,
\end{itemize}
\item if $ m = \text{\textit{drop}}$, 
$(q,wa,\mu) \rightarrow_{\ppda} (q',wa, \mu')$,  
where $\mu'$ is the $\{ 1, \ldots , i, i+1 \}$-pebble assignment such that
$\mu'(j) = \mu(j)$, for each
$j \leq i$, and $\mu'(i+1) = wa$, and
\item if $m = \text{\textit{lift}}$, and $i \in S$, i.e. the last pebble dropped belong of the set of pebble we test the presence of,
$(q,wa,\mu) \rightarrow_{\ppda} (q',wa, \mu')$, 
where $\mu'$ is the $\{ 1, \ldots , i - 1 \}$-pebble assignment such that
$\mu'(j) = \mu(j)$, for each
$j < i$.
\end{itemize}
%\end{definition}
\end{samepage}

\par\noindent\ignorespacesafterend
We are interested in games over pebble pushdown automata, % namely, 
%reachability, B\"uchi, and parity games.
mainly, parity games.

Again, given
a 
%labeled
 transition system, one needs only to provide 
%a mapping $\Omega$ and 
a partition of the set of
configurations 
%$S$ into two sets $S_0$ and $S_1$
 to obtain an arena.
 %
Given a partition of
$Q$ into $Q_0$ and $Q_1$, 
we partition the configurations of
$T_{\ppda}$
into
$\Conf_{\ppda,0}=Q_0\times \Gamma^* \times ( \Gamma^* \biguplus \{ \bot \})^{\{1, \ldots n\}}$
and
$\Conf_{\ppda,1}=Q_1\times  \Gamma^* \times ( \Gamma^* \biguplus \{ \bot \})^{\{1, \ldots n\}}$.

With these notations in mind one can  define the arena
$$
A_{(\ppda, Q_0 , Q_1)}
=
(\Conf_{\ppda,0}, \Conf_{\ppda,1}, \rightarrow_{\ppda}).$$

As expected, given a 
priority function $\Omega: Q \to \{0, \ldots m\}$,
one naturally 
set the extension of $\Omega$ as
$\overline{\Omega}: \Conf(\ppda) \to \{0, \ldots m\}$
and
$\overline{\Omega}(q, w,\mu) = \Omega(q)$
for all
$w \in  \Gamma^*$
and
$\mu \in ( \Gamma^* \biguplus \{ \bot \})^{\{1, \ldots, n\}}$. \\

Concerning pebble pushdown automata, we are interested in the following decision problem.

\problemx{$n$-pebble pushdown parity game}
{An $n$-pebble pushdown automaton $\ppda= (Q, \Gamma,  R, q_{init}, {\gamma_{init}} , F)$, 
where $Q = Q_0  \biguplus Q_1 $, 
and
a priority mapping $\Omega: Q \to \{0, \ldots m\}$.}
{Does player $0$ have a winning strategy from $q_{init}(\gamma_{init}, \mu_{init})$ for the parity game 
$\mathcal{G}= (A_{(\ppda, Q_0 , Q_1)}, \win_{\overline{\Omega}})$ ?\newline}

% \subsubsection{Reduction from parameters to pebble pushdown automata}
\subsection{From parametric pushdown 
automata 
to pebble pushdown automata
}

We now provide a reduction from the problem of solving parametric pushdown parity games to the problem of solving pebble pushdown parity games.

\begin{theorem}\label{PPDA reduction}
%The problem of 
{\sc $n$-parametric pushdown parity game} 
is polynomial time reducible to
%the problem of 
{\sc $n$-pebble pushdown parity game}.
\end{theorem}

\begin{proof}[Sketch]

Let us fix some
$n$-parametric pushdown automaton $\mathcal{Z}= (Q, \Gamma, 
P%\{p_1, \ldots, p_n \}
, q_{init}, { \gamma_{init}} , F)$,
 disjoint
union $Q = Q_0  \biguplus Q_1$
and $P = P_0 \biguplus P_1$,
and
some priority function 
$\Omega : Q \times \Gamma^* \times ( \Gamma^* \biguplus \{ \bot \})^R \to [0, m ]$.

We construct
% polynomial time
  an 
 $n$-pebble pushdown automaton $\ppda =(Q', \Gamma,  R', q_{init}', {\gamma_{init}} , F')$,
a 
 disjoint
union $Q' = Q_0'  \biguplus Q_1'$
and a mapping $\Omega' : Q \times  \Gamma'^* \times ( \Gamma'^* \biguplus \{ \bot \})^{\{1, \ldots, n\}} \to \{1, \ldots, m \}$,
such that
 %:
%
player $0$ has a winning strategy from 
$(q'_{init},\gamma_{init}, \mu_{init})$
in
		\iffalse
		for the parity condition on
		$\mathcal{G}_{(\ppda, Q'_0, Q'_1, \Omega')}$.
		\fi
the parity game
$\mathcal{G'}= (A_{(\ppda, Q'_0, Q'_1)}, \win_{\overline{\Omega'}})$
if and only if
player $0$ has a 
		\iffalse
		strategy $\sigma_0$ for the parameter valuation game  such that
		for all player $1$ strategies $\sigma_1$ for the parameter valuation game,
		player $0$ have a
		winning strategy from $q_{init}(\gamma_{init})$ for the parity condition on
		$\mathcal{G}_{(\mathcal{Z}, \mu_{\sigma_0, \sigma_1}, P_0, P_1, Q_0 , Q_1, \Omega)}$. 
		\fi
winning strategy from $\mu_\bot$ in 
the parity game
$\mathcal{G} =(A_{(\mathcal{Z}, P_0, P_1, Q_0 , Q_1)}, \win_{\widehat{\Omega}})$,
where		 
$A_{(\mathcal{Z}, P_0, P_1, Q_0 , Q_1)}$
and
$\widehat{\Omega}$
correspond to the definitions on page~\pageref{PPDA reachability game}.

%
%
%
%

The transition graph of the $n$-pebble pushdown automaton $\ppda$ will simulate the possible transitions graphs 
(one for every 
%total 
parameter assignment)
 of the pushdown automaton with $n$ parameters by using pebbles to represent the parameters. 
The $n$-pebble pushdown automaton will first simulate the 
parameter valuation arena.
Players take turns chosing particular stack contents on which to place pebbles,
using a gadget similar as the one from Figure~\ref{quantifiers}. 
Once every parameter has been assigned with the dropping of a corresponding pebble, 
the pebble assignment $\mu$ is fixed.
%%
%
%
%
%
%
%

Further configurations in the $n$-pebble pushdown automaton then consist of a word representing the status of the stack, a state, and a fixed pebble assignment. For a fixed pebble assignment there is then a one for one correspondance between configurations of the $n$-parametric pushdown 
automaton 
%transition system with assignment $\mu$
and these of the $n$-pebble pushdown automaton. The set of states of the $n$-pebble pushdown automaton apart from the initial gadget is the same as 
the set of states of the $n$-parametric pushdown automaton, and 
so are player $0$ states, player $1$ states and the priority mapping.
\end{proof}

\section{Reduction to higher-order pushdown automata}\label{section HPDA}

Higher-order pushdown automata (HPDA for short) were introduced as a 
generalization of pushdown automata \cite{aho1969nested,Gre70,maslov1976multilevel}. 
% Alfred V. Aho. Nested stack automata. J. ACM, 16(3):383–406, 1969.
% S. Greibach. Full AFL’s and nested iterated substitution. Inf. Control, 16(1):7–35, 1970.
% A.N. Maslov. Multilevel stack automata. Problemy Peredachi Informatsii, 12:55–62, 1976.
A stack of a pushdown automaton is seen as a {\em level $1$ stack}. 
A pushdown automaton of level $2$ (or $2$-HPDA) then works with a stack of level $1$ stacks. 
In addition to the ability to push
and to pop a symbol on the top-most level $1$ stack, an $2$-HPDA can copy or remove the entire topmost level $1$ stack. 
The definition generalizes to any $n \geq 2$, and $n$-HPDA are similarly defined for all level
$n$ as automata working with a stack of level $(n-1)$ stacks.
% HPDA have been extensively studied as language recognizers [Dam82,Eng83].

We recall the definition from \cite{cachat2007complexity} which itself is taken from \cite{knapik2002higher}. We then provide a
reduction from the problem of solving pebble pushdown parity games to the problem of solving
higher-order pushdown parity games.

\subsection{Higher-order pushdown automata}

A {\em level $1$ stack} (or {\em $1$-stack}) over an alphabet $\Gamma$ is 
simply a %word in $\Gamma^*$
	stack over $\Gamma$, i.e. a word  in $\Gamma^*$.
% an arbitrary sequence $w_0 w_1 \ldots w_n $ of elements of $\Gamma$ for some $n \in \N$. 
A  {\em level $n$ stack} (or {\em $n$-stack}) over an alphabet $\Gamma$, for $n \geq 2$, is a 
non-empty sequence
$ \langle s_0 \rangle \langle s_1 \rangle \ldots \langle s_m \rangle$
of
$(n-1)$-stacks over $\Gamma$,
for some $m \in \N$.
The set of $n$-stacks over $\Gamma$ is denoted by $\mathscr{S}_n(\Gamma)$,
or simply $\mathscr{S}_n$ in case the set $\Gamma$ is obvious from context.
The set of all stacks over $\Gamma$ is written $\mathscr{S}(\Gamma) = \bigcup_{n\in \N} \mathscr{S}_n(\Gamma)$.
We define $\epsilon_{1}$ as
$\epsilon \in \Gamma^*$ and we inductively define 
$\epsilon_{n} = \langle \epsilon_{n-1} \rangle$ in $\mathscr{S}_n$ for all $n > 1$.

A {\em higher-order stack operation} is a partial function from $\mathscr{S}(\Gamma)$ 
to $\mathscr{S}(\Gamma)$ which preserves the level of the input (i.e. the image of
an $n$-stack is an $n$-stack for all $n \in \N$). The {\em level } of an operation $op$
is the smallest $n \in \N$ such that $\text{Dom}(op) \cap \mathscr{S}_n \neq \emptyset$. 
The operations 
%we consider
additionally
 respect the hierarchicality of higher-order stacks, i.e. in
a level $n+1$ stack only the topmost level $n$ stack can be accessed. An operation
$op$ of level $n$, applied to a level $n+1$ stack 
$ \langle s_0 \rangle \langle s_1 \rangle \ldots \langle s_m \rangle$
of length $m \in \N$,
thus
returns the output
$ \langle s_0 \rangle \langle s_1 \rangle \ldots \langle op(s_m) \rangle$
if applicable.
The definition for  all levels of stacks greater than $n$ follows the same pattern.

% For all $m \in \N$, 
The following operations can be performed on a $1$-stacks of length $m \in \N$.
% $ w_0 w_1 \ldots w_m \in \Gamma^*$.
\begin{eqnarray*}
&\text{\textit{push}}_1^\gamma( w_0 w_1 \ldots w_m) &=\quad w_0 w_1 \ldots w_m \gamma   \text{ for all } \gamma \in  \Gamma, \\
&\text{\textit{pop}}_1(w_0 w_1 \ldots w_{m-1} w_m ) &=\quad w_0 w_1 \ldots w_{m-1}, \text{ if } m \geq 1\\
&\text{\textit{top}}( w_0 w_1 \ldots w_m ) &=\quad w_m.
\end{eqnarray*}
The operations added at level $n+1$ are the copy of the topmost $n$-stack
and the removal of the topmost $n$-stack.
More formally, if $ \langle s_0 \rangle \langle s_1 \rangle \ldots \langle s_m \rangle$ is a stack of 
level $n > 1$, 
the following operations are possible.
%
\begin{eqnarray*}
&\text{\textit{push}}_n( \langle s_0 \rangle \langle s_1 \rangle \ldots \langle s_m \rangle) &=\quad  \langle s_0 \rangle \langle s_1 \rangle \ldots \langle s_m \rangle \langle s_m \rangle , \\
&\text{\textit{push}}_j( \langle s_0 \rangle \langle s_1 \rangle \ldots \langle s_m \rangle) &=\quad \langle s_0 \rangle, \langle s_1 \rangle \ldots 
\langle \text{\textit{push}}_j(s_m) \rangle,
\text{ if } 2 \leq j < n\\
&\text{\textit{push}}_1^\gamma( \langle s_0 \rangle \langle s_1 \rangle \ldots \langle s_m \rangle) &= \quad
\langle s_0 \rangle \langle s_1 \rangle \ldots 
\langle \text{\textit{push}}_1^\gamma(s_m) \rangle,
\text{ for all } \gamma \in \Gamma,\\
&\text{\textit{pop}}_n( \langle s_0 \rangle \langle s_1 \rangle \ldots \langle s_{m-1} \rangle \langle s_m \rangle) &= \quad
\langle s_0 \rangle \langle s_1 \rangle \ldots \langle s_{m-1} \rangle
,\\
&\text{\textit{pop}}_j( \langle s_0 \rangle \langle s_1 \rangle \ldots \langle s_{m-1} \rangle \langle s_m \rangle) &= \quad
\langle s_0 \rangle \langle s_1 \rangle \ldots \langle s_{m-1} \rangle
\langle \text{\textit{pop}}_j(s_m) \rangle,
\text{ if } 1 \leq j < n\\
&\text{\textit{top}}( \langle s_0 \rangle \langle s_1 \rangle \ldots \langle s_m \rangle) &=\quad
\text{\textit{top}}(s_m).
\end{eqnarray*}
% We omit the stack alphabet $\Gamma$ from 
%The operation $pop_j$ is undefined on a stack whose top stack of level $j$ is empty.
The operations $pop_1$ and $top$ are undefined on a stack whose top $1$-stack is empty.  \\

Given a stack alphabet $\Gamma$ and $n \in \N$, 
we denote by  $\Op_n(\Gamma)$ the {\em base set of $n$-stack operations} as

\begin{itemize}
\item $\text{\textit{push}}_j$ for all $2 \leq j \leq n$,

\item $\text{\textit{push}}_1^\gamma$ for all $\gamma \in \Gamma$,

\item $\text{\textit{pop}}_j$ for all $1 \leq j \leq n$,

\item $\text{\textit{skip}}$, corresponding to the identity function of $\mathscr{S}(\Gamma)$.

\end{itemize}

% \begin{definition}
\begin{samepage}
\par\noindent\ignorespacesafterend
A {\em higher-order pushdown automata of level $n$} (or $n$-HPDA for short) 
is a tuple $\H=(Q, \Gamma, R, q_{init}, \gamma_{init}, F)$,
where 
\begin{itemize}
        \item $Q$ is a non-empty finite {\em set of  states},
        \item $\Gamma$ is a non-empty finite {\em  stack alphabet},
	\item $R\subseteq Q\times \Gamma \times Q \times \Op_n(\Gamma)$ is a finite {\em set of  rules},
        \item $q_{init}$ is the {\em initial  state},
        \item $\gamma_{init} \in \Gamma$ is the {\em %bottom-of-stack
initial stack symbol}, and
	\item $F\subseteq Q$ is a {\em set of final  states}.
		\end{itemize}
\end{samepage}
% \end{definition}

\par\noindent\ignorespacesafterend
By $\Conf(\mathcal{H}) = Q \times \mathscr{S}_n(\Gamma)$
we denote the set of {\em configurations} of an $n$-HPDA $\H$. 
%is a pair $(q, s)$ where $q \in Q$ and $s$ is an $n$-stack over $\Gamma$. 
As usual we abbreviate $(q,s) \in \Conf(\mathcal{H})$ as $q(s)$.

An $n$-HPDA $\H=(Q, \Gamma, R, q_{init}, F)$
induces the transition system $\T_\H = (Conf(\mathcal{H}), \rightarrow_{\mathcal{H}})$
where
for all $q,q'$ in $Q$,
for all $s,s'$ in $\mathscr{S}_n(\Gamma)$,
$q(s) \rightarrow_{\mathcal{H}} q'(s')$ if
there exists some rule
%$(q,\gamma,q',\theta) \in R$
$(q,\gamma,q',op) \in R$
such that
$\text{\textit{top}}(s) = \gamma$
and
%$s' = \theta(s)$.
$s' = op(s)$.

\iffalse
For our constructions it would be simpler to assume that k-HPDS can work
also on stacks of lower levels, in particular on 1-stacks. Of course we can always
simulate a j-stack, for j < k with an k-stack but in the notation it requires some
additional parenthesis that make it less readable.

To define a game on the graph of a HPDA, we assign a player to each control
state, and we consider an initial configuration: a game structure on a HPDS H
is a tuple G = (H, P0 , P1 , s0 ), where P = P0 ⊎ P1 is a partition of the control
states of H, and s0 ∈ Sk . This extends naturally to a partition of the set of
configurations: with the notations of Section 2.1, V0 = P0 × Sk , V1 = P1 × Sk ,
and E is defined above.
\fi

Again we are interested in parity games.
As expected, given an 
 $n$-HPDA $\mathcal{H}$
 and a partition of
$Q$ into $Q_0$ and $Q_1$,
we partition the configurations of
$T_{\mathcal{H}}$
into
$\Conf_{\mathcal{H},0}=Q_0\times \mathscr{S}_n(\Gamma)$
and
$\Conf_{\mathcal{H},1}=Q_1\times\mathscr{S}_n(\Gamma)$.
%
%
With these notations in mind one can define the arena
$$
A_{(\mathcal{H}, Q_0, Q_1)}
=
(\Conf_{\mathcal{H},0}, \Conf_{\mathcal{H},1}, \rightarrow_{\mathcal{H}})$$
\par\noindent\ignorespacesafterend
induced by an $n$-HPDA $\mathcal{H}$ and a partition of its set of states.

As expected, given a 
priority function $\Omega: Q \to [0, m]$,
we naturally extend the function
as follows,
by setting
$\Omega_{\mathscr{S}_n(\Gamma)}: \Conf(\mathcal{H}) \to [0, m]$
and
$\Omega_{\mathscr{S}_n(\Gamma)}(q, s) = \Omega(q)$
for all
$s \in \mathscr{S}_n(\Gamma)$.

Concerning higher-order pushdown automata, we are interested in the following decision problem. 
% \newpage

\begin{samepage}
\problemx{$n$-HPDA parity game}
{An $n$-HPDA $\H= (Q, \Gamma,  R, q_{init}, {\gamma_{init}} , F)$, 
where $Q = Q_0  \biguplus Q_1 $, 
and
a priority mapping $\Omega: Q \to \{0, \ldots m\}$.}
{Does player $0$ have a winning strategy 
from $q_{init}(\text{\textit{push}}_1^{\gamma_{init}}(\epsilon_{n}))$ 
for the parity game 
$\mathcal{G}= (A_{(\H, Q_0 , Q_1)}, \win_{\Omega_{\mathscr{S}_n(\Gamma)}})$ ?\newline}
\end{samepage}

It was shown in \cite{Cach03} that {\sc $n$-HPDA parity game} can
be solved in $n$-$\EXP$. This also gives an $n$-$\EXP$ algorithm for the $\mu$-calculus
model checking over transitions systems induced by  $n$-HPDA.
In \cite{cachat2007complexity} the matching lower bound was showed, even in the case of reachability games,
hence showing $n$-$\EXP$-completeness of {\sc $n$-HPDA parity game}.

\begin{theorem}{\cite{ Cach03, cachat2007complexity}}
{\sc $n$-HPDA parity game} is $n$-$\EXP$-complete.
\end{theorem}

% \subsection{From pebble pushdown automata to higher-order pushdown automata}

In a $2$-HPDA, the operation $\text{\textit{push}}_2$ % allowig to copy a part of the stack, are responsible for the fact the hierarchy of HPDS is strict.
allows to ``copy'' the top level $1$ stack. The current word is hereby stored away and left untouched
until the next operation $\text{\textit{pop}}_2$,
while $\text{\textit{push}}_1$ and $\text{\textit{pop}}_1$ can be performed on the additional ``copy'' in the meantime. This behavior is similar to that of dropping and lifting a pebble. 
The main difference is that there is no operation to test that the ``copy'', after many updates, is again equal to the ``original'', i.e. there is no operation to syntactically test that the two topmost level $1$ stacks are identical.

% But cannot test

\iffalse
In an $n$-HPDA, the operation $\text{\textit{push}}_j$ % allowig to copy a part of the stack, are responsible for the fact the hierarchy of HPDS is strict.
allows to ``copy'' the top stack of level $j$ for all $2 \leq j \leq k$. 
\fi

In \cite{CaWoe03, Woeh05, carayol2006automates} however 
%the authors
Carayol and W\"ohrle
 introduced a variant of $n$-HPDA by extending the 
$\text{\textit{pop}}_j$ operations for $2 \leq j \leq n$ with a built-in equality test. 
For $2 \leq j \leq n$ 
the new operation
$\text{\textit{pop}}_j^=$ has the same effect as $\text{\textit{pop}}_j$, but
can only be applied if the two top level $j$ stacks coincide.
In \cite{carayol2006automates} it is seen as a symmetrical operation in comparison to
$\text{\textit{push}}_j$.\\

A {\em higher-order pushdown automaton with equality pop of level $n$} ($n$-HPDA$^=$ for short) 
is a higher-order pushdown automaton of level $n$ where in
% Instrn
$\Op_n(\Gamma)$
 the operation 
 %popk
 $\text{\textit{pop}}_j$  
  is replaced by 
%  pop=
%k 
$\text{\textit{push}}_j^=$
%for 1 ≤ k ≤ n. 
for $2 \leq j \leq n$.
We denote
this new set of operations by 
% Instr= n
$\Op_n^=(\Gamma)$. 
More formally the new operations $\text{\textit{push}}_j^=$ are defined as
\begin{eqnarray*}
&\text{\textit{pop}}_k^=( \langle s_0 \rangle \langle s_1 \rangle \ldots  \langle s_m \rangle \langle s_m \rangle) &=\quad  \langle s_0 \rangle \langle s_1 \rangle \ldots \langle s_m \rangle , \quad \text{ and }\\
&\text{\textit{pop}}_j^=( \langle s_0 \rangle \langle s_1 \rangle \ldots \langle s_m \rangle) &=\quad \langle s_0 \rangle \langle s_1 \rangle \ldots 
\langle \text{\textit{pop}}_j^=(s_m) \rangle,
\text{ if } 2 \leq j < k. 
\end{eqnarray*}
%
%
%
%
\iffalse
Dans ce sous-paragraphe, nous présentons formellement les automates à pile
de niveau k sur le jeu d'opérations symétriques Opsk et sur le jeu d'opérations
classiques COpsk . Nous donnons à cette occasion un bref historique de la notion
d'automate à pile de piles. Enfin, nous établissons l'équivalence entre les deux
modèles d'automates à chaque niveau en tant qu'accepteurs de langages. Ce résultat a été obtenu dans [CW03] en collaboration ave Stefan Wöhrle. La preuve
complète apparait dans [Wöh05]. Indépendamment, ce résultat a été obtenu dans
[Fra05].
\fi
%
Transition systems induced by higher-order pushdown automata with equality pop of level $n$
are then defined as expected.  
Carayol and W\"ohrle proved \cite{Woeh05, carayol2006automates} that the two models,
namely
higher-order pushdown automata with equality pop of level $n$
and
higher-order pushdown automata of level $n$,
generate the same classes of transition systems.

\begin{theorem}{\cite{Woeh05, carayol2006automates}}
If $\H$ is an $n$-HPDA (resp. $n$-HPDA$^=$) then there
exists an $n$-HPDA$^=$ (resp. $n$-HPDA) $\H'$ such that
$\H$ and $\H'$ %generate the same graph.
induce isomorphic transition systems.
\end{theorem}

\iffalse
\begin{theorem}{\cite{Woeh05, carayol2006automates}}
If $\H$ is an $n$-HPDA  then there
exists an $n$-HPDA$^=$  $\H'$ such that
$\H$ and $\H'$ %generate the same graph.
induce isomorphic transition systems.
\end{theorem}
% Proposition 3.12

\begin{theorem}{\cite{Woeh05, carayol2006automates}}
If $\H$ is an $n$-HPDA$^=$ then there
exists an $n$-HPDA $\H'$ such that
$\H$ and $\H'$ %generate the same graph.
induce isomorphic transition systems.
\end{theorem}
